The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2023-02-06 SI.tex
Add yaqd-control user guide to SI

@@

@@

>>>>

\RequirePackage{etoolbox}
\RequirePackage{pmboxdraw}
\RequirePackage[section]{minted}
\usemintedstyle{default}
\setminted[python]{linenos=false, frame=none}

\definecolor{bg}{rgb}{0.95, 0.95, 0.95}

\RequirePackage[many]{tcolorbox}
\tcbuselibrary{minted}

@@

\NewTCBListing[number within=chapter, auto counter]{codefragment}{mo}{%
  %colback=bg,
  boxrule=1pt,
  %colframe=bg,
  arc=0pt,
  shadow=false,
  new/use counter=equation,
  boxsep=1ex, top=0pt, left=0pt, right=0pt, bottom=0pt,
  comment={\hfill(\arabic{chapter}.\arabic{equation})},
  listing outside comment,
  righthand width=3em,
  sidebyside gap=0pt,
  minted language=#1,
  %before skip =-0.5\baselinestretch,
  %after skip=2\baselinestretch,
}

\BeforeBeginEnvironment{codefragment}{\begin{singlespace}\stepcounter{equation}}
\AfterEndEnvironment{codefragment}{\end{singlespace}}

\newmintinline[bash]{bash}{bgcolor=bg}
\newmintinline[python]{python}{bgcolor=bg}

@@

\section{\texttt{yaqd-control}: tooling for managing \yaq{} daemons}

\texttt{yaqd-control} is a command line utility provided by the \yaq{} project to manage \yaq{} daemons.
The primary purpose of the tool is to maintain a list of active daemons, provide access to their configuration, and their running state.
This includes querying the active state of known daemons, as well as starting daemons running as backgound services.

\subsection{installation}

yaqd-control can be installed via
PyPI\cite{yaqd-control} or
conda-forge\cite{yaqd-control-conda}.

\begin{codefragment}{bash}
$ pip install yaqd-control
\end{codefragment}

\begin{codefragment}{bash}
$ conda config --add channels conda-forge
$ conda install yaqd-control
\end{codefragment}

\subsection{Usage}

yaqd-control is a command line application.

Help: learn more, right from your terminal.

\begin{codefragment}{bash}
$ yaqd --help
Usage: yaqd [OPTIONS] COMMAND [ARGS]...

Options:
  --help  Show this message and exit.

Commands:
  clear-cache
  disable
  edit-config
  enable
  list
  reload
  restart
  scan
  start
  status
  stop
\end{codefragment}

Try \texttt{yaqd\ \ -\/-help} to learn more about a particular command.

\paragraph{the cache}\label{the-cache}

yaqd-control keeps track of known daemons, referred to as the cache

Status: yaqd-control can quickly show you the status of all daemons in
yaqd-control's cache. This is usually the most used subcommand, as it
gives a quick overview of the system, which daemons are offline, and
which are currently busy.

\begin{codefragment}{bash}
$ yaqd status
+-----------+-------+--------------------------+------+---------+-------+
| host      | port  | kind                     | name | status  | busy  |
+-----------+-------+--------------------------+------+---------+-------+
| 127.0.0.1 | 38202 | system-monitor           | foo  | online  | False |
| 127.0.0.1 | 39054 | fake-continuous-hardware | bar  | online  | True  |
| 127.0.0.1 | 39055 | fake-continuous-hardware | baz  | online  | False |
| 127.0.0.1 | 39056 | fake-continuous-hardware | spam | offline | ?     |
| 127.0.0.1 | 37067 | fake-discrete-hardware   | ham  | online  | False |
| 127.0.0.1 | 37066 | fake-discrete-hardware   | eggs | online  | False |
+-----------+-------+--------------------------+------+---------+-------+
\end{codefragment}

List: this is essentially the same as \texttt{status} except that it
does not attempt to contact the daemons, so it does not give you
additional context. List supports a flag \texttt{--format} which accepts "json",
"toml", "prettytable", or "happi".

\begin{codefragment}{bash}
$ yaqd list
+-----------+-------+--------------------------+------+
| host      | port  | kind                     | name |
+-----------+-------+--------------------------+------+
| 127.0.0.1 | 38202 | system-monitor           | foo  |
| 127.0.0.1 | 39054 | fake-continuous-hardware | bar  |
| 127.0.0.1 | 39055 | fake-continuous-hardware | baz  |
| 127.0.0.1 | 39056 | fake-continuous-hardware | spam |
| 127.0.0.1 | 37067 | fake-discrete-hardware   | ham  |
| 127.0.0.1 | 37066 | fake-discrete-hardware   | eggs |
+-----------+-------+--------------------------+------+
\end{codefragment}

Scan: Scanning allows you to add currently running daemons to the cache.

\begin{codefragment}{bash}
$ yaqd scan
scanning host 127.0.0.1 from 36000 to 39999...
...saw unchanged daemon fake-discrete-hardware:eggs on port 37066
...saw unchanged daemon fake-discrete-hardware:ham on port 37067
...found new daemon system-monitor:foo on port 38202
...found new daemon fake-continuous-hardware:bar on port 39054
...saw unchanged daemon fake-continuous-hardware:baz on port 39055
...known daemon fake-continuous-hardware:spam on port 39056 not responding
...done!
\end{codefragment}

Scan has some additional options, passed as flags on the command line,
which allow you to change the default scan range and host (for remotely
accessed daemons):

\begin{codefragment}{bash}
$ yaqd scan --help
Usage: yaqd scan [OPTIONS]

Options:
  --host TEXT      Host to scan.
  --start INTEGER  Scan starting point.
  --stop INTEGER   Scan stopping point.
  --help           Show this message and exit.
\end{codefragment}

Edit Config: yaqd-control provides an easy way to edit the default
config file location for a daemon kind. This uses your default editor
(\texttt{EDITOR} environment variable), and defaults to \texttt{notepad.exe} on
Windows, and \texttt{vi} on other platforms. Using yaqd-control to edit
config files means that you do not need to know the default location.
Additionally, it does some basic validity checks (that the toml parses
and that each daemon section has the \texttt{port} keyword). If an error
is found, you are prompted to re-edit the file. Daemons from the config
file are added to the cache. You may pass multiple daemon kinds, which
will be opened in succession.

\begin{codefragment}{bash}
$ yaqd edit-config fake-continuous-hardware system-monitor
\end{codefragment}

Clear Cache: Note that this is a destructive action.
\texttt{clear-cache} deletes all daemons from the cache (thus
\texttt{list} and \texttt{status} will give empty tables) There is no
user feedback.

\begin{codefragment}{bash}
$ yaqd clear-cache
$ yaqd status
+------+------+------+------+--------+------+
| host | port | kind | name | status | busy |
+------+------+------+------+--------+------+
+------+------+------+------+--------+------+
\end{codefragment}

\paragraph{Running in the background}\label{running-in-the-background}
Each of the commands in this section can take multiple daemon kinds.
Enable: by enabling a daemon, you allow the operating system to manage
that daemon in the background. An enabled daemon will always start again
when you restart your computer. Enabling is required for the rest of the
commands in this section to work as expected. After enabling, it's
typical to start the daemon as well, this does not happen automatically.
Enablement works in slightly different ways on different platforms, but
the commands are the same (don't worry if the password prompts are
different). Currently supported platforms are Linux (systemd), MacOS
(launchd) and Windows (via NSSM, bundled with the distribution).

\begin{codefragment}{bash}
$ yaqd enable system-monitor
[sudo] password for scipy2020:
==== AUTHENTICATING FOR org.freedesktop.systemd1.manage-unit-files ===
Authentication is required to manage system service or unit files.
Password:
==== AUTHENTICATION COMPLETE ===
\end{codefragment}

Disable: this is the inverse operation to enable, which makes it so that
the daemon does not start on reboot. This does not affect the running
daemon.

\begin{codefragment}{bash}
$ yaqd disable system-monitor
==== AUTHENTICATING FOR org.freedesktop.systemd1.manage-unit-files ===
Authentication is required to manage system service or unit files.
Password:
==== AUTHENTICATION COMPLETE ===
Removed /etc/systemd/system/multi-user.target.wants/yaqd-system-monitor.service.
\end{codefragment}

Start: This starts the daemon running in the background immediately. It
must have been enabled to run in the background using this command.

\begin{codefragment}{bash}
$ yaqd start system-monitor
==== AUTHENTICATING FOR org.freedesktop.systemd1.manage-units ===
Authentication is required to start 'yaqd-system-monitor.service'.
Password:
==== AUTHENTICATION COMPLETE ===
\end{codefragment}

Stop: This stops the daemon running in the background immediately. It
must have been running in the background using yaqd-control (either on
startup via enable or via the start command above).

\begin{codefragment}{bash}
$ yaqd stop system-monitor
==== AUTHENTICATING FOR org.freedesktop.systemd1.manage-units ===
Authentication is required to stop 'yaqd-system-monitor.service'.
Password:
==== AUTHENTICATION COMPLETE ===
\end{codefragment}

Restart/Reload: This stops (if running) and restarts the daemon running
in the background immediately. Reload is slightly different in that it
signals to the daemon to reload its configuration rather than completely
restart, but effectively it is the same as restart (and is a pure alias
where such a signal is not supported). It must have been enabled to run
in the background using this command.

\begin{codefragment}{bash}
$ yaqd restart system-monitor
==== AUTHENTICATING FOR org.freedesktop.systemd1.manage-units ===
Authentication is required to restart 'yaqd-system-monitor.service'.
Password:
==== AUTHENTICATION COMPLETE ===
\end{codefragment}
